Automatic Matching of Bullet Lands

### 2016-01-17 imaging-paper.Rnw
reference added to stable region

Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
\hh{XXX Note that the definition given in this paper does not coincide with the one given in \citet{thompson:2013}. There, CMS is defined only in terms of matching peaks. Additionally, \citet{thompson:2013} peaks are only used if they can be identified and matched `within a tolerable range' between lands. The definition given here is computationally less complex, but should yield similar values, because of the requirement to only consider signatures from a stable region in the land. In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92). XXX}
For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match. \hh{XXX Biasotti's four or more consecutive lines is closer to eight or more with our definition. Should we just multiply our CMS by 0.5, to account for the fact that between every two peaks has to be a valley? XXX} We will assess this claim in the next section.

@@

\subsection{Impact of cross sections}
The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of which cross section level was used in a matching. Visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion).
However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. 
This situation is assessed and discussed in more detail below:

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
             lop[[4]], lop[[5]], lop[[6]],
             ncol=2)
@
\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet~1-1. Initially, the match between signatures from cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet. At a  level of $175\mu m$ the bullet's signature stabilizes. For this land, matches should not be attempted at lower cross sections. }
\end{figure}
%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
by comparing signatures from cross sections that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures. In the approach here, we keep increasing the height $x$ of the cross section until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.

@@

aucs <- plyr::ldply(features, function(x) {
  subbstats <- subset(bstats, flagged == FALSE)
  pos.scores <- sample(subbstats[which(subbstats$match),x], 5000, replace=TRUE)
  neg.scores <- sample(subbstats[which(!subbstats$match),x], 5000, replace=TRUE)
  data.frame(x, upper=mean(pos.scores > neg.scores), 
             lower = mean(pos.scores < neg.scores))
})
aucs <- melt(aucs, measure.var=c("lower", "upper"))
names(aucs) <- c("variable", "type", "auc")
rocs <- merge(rocs, aucs, by=c("variable", "type"))
rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")

>>>>

Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
\hh{XXX Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}. There, CMS is defined only in terms of matching peaks. Additionally, peaks in  \citet{thompson:2013} are only used if they can be identified and matched `within a tolerable range' between lands. The definition given here is computationally less complex, but should yield similar values, because of the requirement to only consider signatures from a stable region in the land (see Section~\ref{sec:heights} for further details on stability of regions). In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92). XXX}
For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match. \hh{XXX Biasotti's four or more consecutive lines is closer to eight or more with our definition. Should we just multiply our CMS by 0.5, to account for the fact that between every two peaks has to be a valley? XXX} We will assess this claim in the next section.

@@

\subsection{Impact of cross sections}\label{sec:heights}
The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of which cross section level was used in a matching. Visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion).
However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. 
This situation is assessed and discussed in more detail below:

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
             lop[[4]], lop[[5]], lop[[6]],
             ncol=2)
@
\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet~1-1. Initially, the match between signatures from cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet. At a  level of $175\mu m$ the bullet's signature stabilizes. For this land, matches should not be attempted at lower cross sections. }
\end{figure}
%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
by comparing signatures from cross sections that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures. In the approach here, we keep increasing the height $x$ of the cross section until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.

@@

set.seed(20140105)
aucs <- plyr::ldply(features, function(x) {
  subbstats <- subset(bstats, flagged == FALSE)
  pos.scores <- sample(subbstats[which(subbstats$match),x], 10000, replace=TRUE)
  neg.scores <- sample(subbstats[which(!subbstats$match),x], 10000, replace=TRUE)
  data.frame(x, upper=mean(pos.scores > neg.scores), 
             lower = mean(pos.scores < neg.scores))
})
aucs <- melt(aucs, measure.var=c("lower", "upper"))
names(aucs) <- c("variable", "type", "auc")
rocs <- merge(rocs, aucs, by=c("variable", "type"))
rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")
